Update and evaluation of the ozone dry deposition in the Oslo~CTM3 v1.0

### 2019-07-11 DryDep.tex
Corrected order in supplement and manuscrit accordingly.

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $R^\chem{O_3}$ & Emissions\\
    & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & & (year)\\
    \middlehline
    Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
    mOSaic            &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offLight  &   x & 1 & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_SWVL1     &   x & x & VL1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
    mOSaic\_desert    &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{AE:Gusten1995}$^*$  & 2005 \\
    mOSaic\_emis2014  &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2014 \\
    mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.7} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig04}
  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition. The ozone burden increases globally except for some regions covered by tropical forest. Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}. We come back to this in the following sections.

@@

Dry deposition velocities are directly available only for the new model version. For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(p_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$ 
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(p_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
  \label{eq:retro_vdd}
\end{equation}
Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$. In case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly ailable from model output.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$). We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance. Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly. Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}) amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$). Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH. The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.7). The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, where an increase in ozone precursors is seen in CEDS.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b. The dry deposition velocities in the mOSaic scheme are well below the Wesely scheme and in remarkable agreement with the results shown by \citet{ACP:Hardacre2015}. In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean. This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as new standard for the Oslo~CTM3. This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation. The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$. The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.6: Fig.~S8). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$ (almost entirely covered by ocean), $(12-30)\,\unit{^\circ S}$ (coinciding with the location of Australia and its desert regions), as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.

>>>>

\begin{table*}[t]
  \caption{Summary of specifications of all simulations discussed in this section. For simplicity, only the tested parameters are listed. An x denotes that the model was run exactly in the configuration as has been described in Section~\ref{sec:model_des}.}
  \begin{tabular}{lccccccccc}
    \tophline
    \multirow{2}{*}{Simulation} & \multicolumn{3}{c}{mOSaic scheme} & \multicolumn{2}{c}{Greening season} & $R^\chem{O_3}$ & Emissions\\
    & $f_\text{phen}$ & $f_\text{light}$ & $f_\text{SW}$ & SMOKE-BEIS & $5\,\unit{^\circ C}$-days & & (year)\\
    \middlehline
    Wesely\_type      &  \multicolumn{3}{c}{n/a}   & x & n/a& \citet{AE:Wesely1989, JGR:Hough1991}  & 2005 \\
    mOSaic            &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offLight  &   x & 1 & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_offPhen   &   1 & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_SWVL1     &   x & x & VL1              &   &  x & \citet{ACP:Simpson2012}  & 2005 \\
    mOSaic\_ice       &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{ACP:Helmig2007}$^\dagger$ & 2005 \\
    mOSaic\_desert    &   x & x & x                &   &  x & \citet{ACP:Simpson2012}; \citet{AE:Gusten1995}$^*$  & 2005 \\
    mOSaic\_emis2014  &   x & x & x                &   &  x & \citet{ACP:Simpson2012}  & 2014 \\
    mOSaic\_hough     &   x & x & x                &   &  x & \citet{AE:Wesely1989, JGR:Hough1991}$^\star$  & 2005 \\
    \bottomhline
  \end{tabular}
  \label{tab:simsum}
  \belowtable{$^\dagger\,R^\chem{O_3}_\text{ice/snow}=10000\,\unit{s\,m^{-1}}$; $^*\,R^\chem{O_3}_\text{desert}=800\,\unit{s\,m^{-1}}$; $^\star\,$For adapted values see Supplement~S.6} % Table Footnotes
\end{table*}
%
\begin{figure}[t]
  \includegraphics[width=8.3cm]{fig04}
  \caption{Relative difference between reference simulations \emph{mOSaic} and the \emph{Wesely\_type} with respect to (a) average ozone burden in the lowermost model level; (b) average ozone dry deposition velocity; (c) total amount of ozone removed from the atmosphere by dry deposition.}
  \label{fig:diff_maps}
\end{figure}
%
In Fig.~\ref{fig:diff_maps}, we show the average relative difference between \emph{mOSaic} and \emph{Wesely\_type} on global maps by means of ozone burden in the lowermost model level, dry deposition velocity and total ozone dry deposition. The ozone burden increases globally except for some regions covered by tropical forest. Especially in desert regions in Africa, America, and Asia, ozone burden increases by more then $100\,\unit{\%}$. Consistently, dry deposition velocities decrease globally by the same order of magnitude in these regions, while they increase over tropical forest. With respect to total dry deposition, the picture is a bit less clear. We find a decrease of total dry deposition of ozone in desert regions and ocean covered areas and an increase in regions covered by tropical forest, while at mid and high latitudes in both hemispheres only small changes are visible. A possible explanation to this divergence especially in desert regions is the difference between the prescribed surface resistances $R^\chem{O_3}$ in the Wesely scheme in comparison to those used in \emph{mOSaic}. We come back to this in the following sections.

@@

Dry deposition velocities are directly available only for the new model version. For \emph{Wesely\_type}, monthly averaged dry deposition velocities $v^\chem{O_3}_\text{DD}$ had to be retrospectively estimated from the ratio between the total ozone dry deposition $\sum \chem{O_3^\text{DD}}(p_0)$ and monthly averaged ozone amount in the lowermost model level $\chem{O_3}(p_0)$ 
\begin{equation}
  v^\chem{O_3}_\text{DD} = \frac{\sum \chem{O_3^\text{DD}}(p_0)}{\chem{O_3}(p_0)}\cdot c_\text{month}.
  \label{eq:retro_vdd}
\end{equation}
Herein, $c_\text{month} = \frac{\Delta h_\text{month}}{s_\text{month}}$, with the monthly average height of the lowermost model level in each gridbox $\Delta h_\text{month}$ and the respective number of seconds in a month $s_\text{month}$. In case of \emph{mOSaic}, resulting values for $v^\chem{O_3}_\text{DD}$ from Eq.~(\ref{eq:retro_vdd}) are compatible with the values which are directly ailable from model output.
%
\subsubsection{Zonal distribution}
\label{subsubsec:zonal}
%
The annual zonal average with respect to surface ozone concentration (Fig.~\ref{fig:mmm_drydep}a) displays on average, in consistency with Fig.~\ref{fig:diff_maps}a, a global increase of surface ozone concentrations by $6\,\unit{ppb}$ comparing \emph{mOSaic} to \emph{Wesely\_type}. This increase is largest in the zonal band $(25-50)\,\unit{^\circ N}$ which contains the major deserts. In the deep tropics ($5\,\unit{^\circ S}-5\,\unit{^\circ N}$), the increase is smallest ($\mathcal{O}(5\,\unit{ppb})$). We find that the mOSaic scheme further intensifies the strong asymmetry between northern and southern hemisphere as a consequence of the distribution of the continental land masses and vegetation thereon. Among the sensitivity studies focusing on the stomatal conductance, there is only a low absolute variance. Neglecting the dependence on light in the stomatal conductance formulation (\emph{mOSaic\_offLight}) -- or in other words allowing photosynthesis 24/7 -- decreases the ozone concentration by $1-2\,\unit{ppb}$ in the tropics and NH mid latitudes, while choosing soil water at shallower depths (\emph{mOSaic\_SWVL1}) increases $[\chem{O_3}]$ insignificantly. Rather surprisingly, switching off the phenology completely (\emph{mOSaic\_offPhen}) amounts on average only to small difference ($\mathcal{O}(<1\,\unit{ppb})$). Most remarkable, but expected due to the much smaller prescribed dry deposition velocity over ice and snow, \emph{mOSaic\_ice} displays a doubling of surface ozone in the high Arctics compared to \emph{Wesely\_type} ($\mathcal{O}(20\,\unit{ppb})$) but affects ozone concentrations down to latitudes at about $50\,\unit{^\circ}$ in both hemispheres. Reducing $R^\chem{O_3}_\text{desert}$ by $60\,\unit{\%}$ (\emph{mOSaic\_desert}) a reduction in the order of $1\,\unit{ppb}$ is found mainly limited to the NH. The largest impact on ozone concentrations ($\mathcal{O}(2-5\,\unit{ppb})$) is found for the experiment \emph{mOSaic\_hough} which is closest to \emph{Wesely\_type}, since we used on average the same $R^\chem{O_3}$ (see Supplement~S.6). The scenario of differing emissions (2005 in comparison to 2014 or more specifically \emph{mOSaic} compared to \emph{mOSaic\_emis2014}), yields higher ozone concentrations in the northern hemisphere in 2005 in accordance to a reduction in sulfur and \chem{NO_x} emissions in south east Asia in later years. An opposite tendency is seen for latitudes south of $30\,\unit{^\circ N}$, where an increase in ozone precursors is seen in CEDS.

The $v^\chem{O_3}_\text{DD}$ are shown in Fig.~\ref{fig:mmm_drydep}b. The dry deposition velocities in the mOSaic scheme are well below the Wesely scheme and in remarkable agreement with the results shown by \citet{ACP:Hardacre2015}. In the Arctics, except for \emph{mOSaic\_ice}, all model experiments are slightly above the multi-model-mean. This indicates, that with respect to the other models, the \citet{ACP:Helmig2007} surface resistance above ice and snow should be considered as new standard for the Oslo~CTM3. This may, however, lead to an overcompensation of the current arctic low-bias in surface ozone in the Oslo~CTM3 and needs further evaluation. The dry deposition velocities are of course independent of the emission scenario, but display a strong sensitivity on $f_\text{light}$, $f_\text{phen}$, and especially on the choice of $R^\chem{O_3}$. The shape of the normalized zonal average dry deposition velocities of the mOSaic scheme are more similar to the multi-model-mean than to \emph{Wesely\_type} (Supplement~S.7: Fig.~S8). The biggest exceptions are the zonal bands $(50-70)\,\unit{^\circ S}$ (almost entirely covered by ocean), $(12-30)\,\unit{^\circ S}$ (coinciding with the location of Australia and its desert regions), as well as its counterpart in the northern hemisphere $(12-30)\,\unit{^\circ N}$.
